Supplemental Information

### 2022-08-02 manuscript/manuscript.Rmd
finish edits prior to submission

# rbbt::bbt_update_bib(
#   path = "manuscript/manuscript.Rmd",
#   ignore = c("R-base"),
#   path_bib = "manuscript/library.json"
# )

@@

Having identified experimental conditions for exponential growth, we next determined the extracellular fluxes of glucose (GLC), lactate (LAC), pyruvate (PYR), and amino acids (**Figs \@ref(fig:m1)I-J**). Flux calculations incorporated changes in cell number, extracellular metabolite concentrations, metabolite degradation rates, and medium evaporation over time [@murphy2013a] (**Figure S1**). Interestingly, while we observed a modest increase in glucose uptake, we found that hypoxia actually decreased lactate efflux (**Figs \@ref(fig:m1)I**). This decrease in lactate efflux occurred despite activation of the HIF-1α transcriptional program as reflected by increased expression of GLUT1 and LDHA. To test if more severe hypoxia would augment glycolysis, we culture cells in 0.2% ambient oxygen (**Fig S2**). Under these conditions, we observed no change in glucose or lactate fluxes, similar to 0.5% oxygen culture. To test if this unexpected response was unique to LFs, we next studied PASMCs under 0.5% oxygen conditions (**Fig S3**). PASMCs grew faster than LFs, and so samples were collected every 12 h for 48 h for these cells. Again, similar to LFs at 0.5% and 0.2% oxygen, we observed no change in glucose uptake and reduced lactate efflux in PASMCs regardless of HIF-1α stabilization. Together, these data suggest that hypoxia uncouples HIF target gene expression and glycolytic flux in proliferating primary cells.

@@

Although the metabolite exchange fluxes for bidirectional reactions tend to be poorly resolved by metabolic flux analysis [@wiechert2007a], two observations are worth highlighting (**Tables S1-S3**). First, consistent with the stable isotope tracing results, the modeled rate of reductive carboxylation through reverse flux by isocitrate dehydrogenase in LFs is low (~4 fmol/cell/h), unchanged by hypoxia, and modestly increased by BAY treatment. By contrast, the rate of reductive carboxylation increases 6-fold in PASMCs in hypoxia, highlighting an important role for this pathway in the metabolic response of PASMCs to decreased oxygen availability (**Fig \@ref(fig:m4)A**).

@@

For metabolomic profiling and quantification of isotopic enrichment for FBP and 3PG, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B. The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*. Resolution was set to 70,000 and the AGC target was 1×10^6^ ions. Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method. For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run. Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas. Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018]. Missing values were imputed using random forest. Samples peak areas were normalized using probabilistic quotient normalization [@dieterle2006]. Differentially regulated metabolites were identified using limma [@ritchie2015a]. Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021] with KEGG metabolite pathways [@kanehisa2000].

@@

RNA was collected from LFs treated for three days ± hypoxia ± BAY as described above. Four biological replicates were analyzed. Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample. Sequences were deposited in the NIH SRA (PRJNA721596). Sequences were mapped to the human GRCh38 primary assembly and counts summarized using Rsubread [@liao2019]. This data is available from the Oldham Lab GitHub repository (https://github.com/oldhamlab/rnaseq.lf.hypoxia.molidustat). Differentially expressed transcripts were identified using DESeq2 [@love2014]. Gene set enrichment and transcription factor enrichment was performed using the fgsea and DoRothEA R packages, respectively [@korotkevich2021; @garcia-alonso2019].

@@

(ref:m1) **Effects of hypoxia on extracellular metabolite fluxes in lung fibroblasts.** (**A**) Lung fibroblasts (LFs) were cultured in 21% (*red*) or 0.5% (*blue*) oxygen beginning 24 h prior to time 0. Samples were collected every 24 h for 72 h. (**B**) Growth curves of LFs in each experimental condition (n = 8). (**C**) Growth rates from (B) were determined by robust linear modeling of log-transformed growth curves. (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to 21% oxygen at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to 21% oxygen treatment at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 8). (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM. Comparisons were made using linear mixed effects models with oxygen as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (* p < 0.05).

(ref:m2) **Effects of pharmacologic prolyl hydroxylase inhibition on extracellular metabolite fluxes in lung fibroblasts.** (**A**) Lung fibroblasts (LFs) were treated with PHD inhibition molidustat (BAY, 10 μM) (*purple*) or DMSO (*green*) beginning 24 h prior to time 0. Samples were collected every 24 h for 72 h. (**B**) Growth curves of LFs in each experimental condition (n = 8). (**C**) Growth rates from (B). (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to DMSO at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to DMSO at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 8). (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (* p < 0.05).

(ref:m3) **Metabolic flux analysis of lung fibroblasts following hypoxic and pharmacologic PHD inhibition.** (**A**) Mass isotopomer distribution (MID) of pyruvate (PYR) following 72 h labeling of LFs with [U-^13^C~6~]-glucose (GLC). (**B**) MID of citrate after 72 h labeling of LFs with [U-^13^C~6~]-GLC (**C**) MID of citrate after 72 h labeling of LFs with [U-^13^C~5~]-glutamine (GLN). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (n = 4, p < 0.05 indicated as * 0.5% *v.* 21% oxygen, † BAY *v.* DMSO, ‡ Δoxygen *v.* ΔBAY). (**D**) Fraction of M5 citrate indicating reductive carboxylation after labeling with [U-^13^C~5~]-GLN in LFs and PASMCs (n = 3-4, * p < 0.05 0.5% *v.* 21% oxygen). (**E**) Ratio of metabolic fluxes in 0.5% oxygen compared to 21% oxygen. Fluxes with non-overlapping confidence intervals are highlighted with arrows colored according to the magnitude of the change. Arrow thickness corresponds to the absolute flux measured in hypoxia. (**F**) Ratio of metabolic fluxes in BAY-treated cells compared to DMSO-treated control. Arrows are colored as in (E) and arrow weights correspond to the absolute flux as measured in BAY-treated cells.

(ref:m4) **PHD inhibition increases lactate uptake and oxidation.** (**A**) Exchange flux estimates for reductive carboxylation of α-ketoglutarate (AKG) to citrate (CIT). Data show the estimate with upper and lower bounds for LFs and PASMCs. (**B**) Exchange flux estimates for lactate import presented as in (A). (**C**) Mass isotopomer distributions of lactate (LAC), fructose-bisphosphate (FBP), pyruvate (PYR), and citrate (CIT) in LFs following 72 h labeling with [U-^13^C~3~] lactate. Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (n = 4, p < 0.05 indicated as * 0.5% *v.* 21% oxygen, † BAY *v.* DMSO, ‡ Δoxygen *v.* ΔBAY).

(ref:m5) **Metabolomic analysis of molidustat treatment in normoxia and hypoxia.** (**A**) LFs were cultured in standard growth medium and treated with molidustat (BAY, 10 μM) or vehicle (DMSO, 0.1%) in 21% or 0.5% oxygen conditions for 72 h. Growth rates were determined by linear fitting of log-transformed growth curves. (**B**) Glucose uptake in LFs treated with BAY cultured in hypoxia (note reversed *y*-axis). (**C**) Lactate efflux in LFs treated with BAY cultured in hypoxia. Data are mean ± SEM. Comparisons were made using linear mixed effects models with oxygen × treatment as the fixed effects and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (* p < 0.05 BAY *v.* DMSO within a given oxygen exposure). (**D**) Principal component analysis of intracellular metabolites following 72 h of treatment as in (A) suggests a dominant effect of hypoxia over pharmacologic PHD inhibition on the metabolome (n = 7). (**E**) Volcano plot of intracellular metabolites of BAY-treated cells cultures in 21% or 0.5% oxygen. *Blue* circles indicate increased levels in hypoxia while *red* circles indicate decreases with hypoxia. (**F**) Metabolite set enrichment analysis of metabolites from (E). KEGG pathways with p-values < 0.05 are shown. (**G**) Leading edge analysis of the TCA cycle metabolite set. Negative values indicate relative enrichment associated with BAY treatment compared to hypoxia treatment. Abbreviations: PYR, pyruvate; SUC, succinate; PEP, phosphoenolpyruvate; CIT, citrate; AKG, α-ketoglutarate; MAL, malate; ACO, aconitate; FUM, fumarate. (**H**) NAD^+^ from LFs cultured as in (A). (**I**) NADH from LFs cultured as in (A). (**J**) NADH/NAD^+^ ratio from (H) and (I). NAD and NADH were determined by enzymatic assay. Data are mean ± SEM (n = 5). Comparisons were made using linear mixed effects models with treatment as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (p < 0.05 with *black* * indicating a significant effect of treatment within a given oxygen tension and *colored* * indicating a significant effect of treatment within a given oxygen tension denoted by the color).

(ref:m6) **Transcriptomic analysis of molidustat treatment in normoxia and hypoxia.** (**A**) Principal component analysis of transcriptional changes following 72 h of treatment with 0.5% oxygen or molidustat (BAY), separately or together (n = 4). (**B**) Volcano plot illustrating the effects of hypoxia in BAY-treated cells on gene expression. (**C**) Gene set enrichment analysis of transcripts from (B). (**D**) Volcano plot of those transcripts comprising the E2F Targets and G2/M Checkpoint Hallmark gene sets. (**E**) Volcano plot of those transcripts comprising the MYC Targets V1 and V2 Hallmark gene sets. (**F**) Volcano plot illustrating the results of a transcription factor enrichment analysis suggests mechanisms for differential regulation of gene expression following hypoxia or BAY treatment.

(ref:m7) **MYC regulates HIF-dependent glycolytic flux.** (**A**) Representative immunoblot of MYC protein expression in lung fibroblasts following 72 h of treatment with 0.5% oxygen or molidustat (BAY, B). (**B**) Quantification of band densities from (A). (**C**) Working model illustrating that hypoxia-induced MYC antagonizes HIF-dependent metabolic and transcriptional events. (**D**) Representative immunoblot of lung fibroblasts (LFs) treated with siRNA targeting MYC (M) demonstrating adequate protein knockdown. (**E**) Growth rates of MYC-knockdown cells cultured in hypoxia. (**F**) Lactate efflux rates of MYC-knockdown cells cultured in hypoxia. (**G**) Representative immunoblot of LFs treated with MYC adenovirus. (**H**) Growth rates of MYC overexpressing cells cultured with BAY. (**I**) Lactate efflux rates of MYC overexpressing cells cultured with BAY. Data are mean ± SEM of n = 3-7 biological replicates. Comparisons were made using linear mixed effects models with oxygen × treatment as the fixed effects and biological replicate as a random effect. The multivariate *t* distribution was used to determine significant differences (p < 0.05) as indicated by black * for differences within groups and colored * for differences between groups defined by the *x*-axis.

>>>>

rbbt::bbt_update_bib(
  path = "manuscript/manuscript.Rmd",
  ignore = c("R-base"),
  path_bib = "manuscript/library.json"
)

@@

Having identified experimental conditions for exponential growth, we next determined the extracellular fluxes of glucose (GLC), lactate (LAC), pyruvate (PYR), and amino acids (**Figs \@ref(fig:m1)I-J**). Flux calculations incorporated changes in cell number, extracellular metabolite concentrations, metabolite degradation rates, and medium evaporation over time [@murphy2013a] (**Figure S1**). Interestingly, while we observed a modest increase in glucose uptake, we found that hypoxia actually decreased lactate efflux (**Figs \@ref(fig:m1)I**). This decrease in lactate efflux occurred despite activation of the HIF-1α transcriptional program as reflected by increased expression of GLUT1 and LDHA. To test if more severe hypoxia would augment glycolysis, we cultured cells in 0.2% ambient oxygen (**Fig S2**). Under these conditions, we observed no change in glucose or lactate fluxes, similar to 0.5% oxygen culture. To test if this unexpected response was unique to LFs, we next studied PASMCs under 0.5% oxygen conditions (**Fig S3**). PASMCs grew faster than LFs, and so samples were collected every 12 h for 48 h for these cells. Again, similar to LFs at 0.5% and 0.2% oxygen, we observed no change in glucose uptake and reduced lactate efflux in PASMCs regardless of HIF-1α stabilization. Together, these data suggest that hypoxia uncouples HIF target gene expression and glycolytic flux in proliferating primary cells.

@@

Although the metabolite exchange fluxes for bidirectional reactions tend to be poorly resolved by metabolic flux analysis [@wiechert2007], two observations are worth highlighting (**Tables S1-S3**). First, consistent with the stable isotope tracing results, the modeled rate of reductive carboxylation through reverse flux by isocitrate dehydrogenase in LFs is low (~4 fmol/cell/h), unchanged by hypoxia, and modestly increased by BAY treatment. By contrast, the rate of reductive carboxylation increases 6-fold in PASMCs in hypoxia, highlighting an important role for this pathway in the metabolic response of PASMCs to decreased oxygen availability (**Fig \@ref(fig:m4)A**).

@@

For metabolomic profiling and quantification of isotopic enrichment for FBP and 3PG, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B. The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*. Resolution was set to 70,000 and the AGC target was 1×10^6^ ions. Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method. For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run. Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas. Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018]. Missing values were imputed using random forest. Samples peak areas were normalized using probabilistic quotient normalization [@dieterle2006]. Differentially regulated metabolites were identified using limma [@ritchie2015a]. Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021a] with KEGG metabolite pathways [@kanehisa2000].

@@

RNA was collected from LFs treated for three days ± hypoxia ± BAY as described above. Four biological replicates were analyzed. Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample. Sequences were deposited in the NIH SRA (PRJNA721596). Sequences were mapped to the human GRCh38 primary assembly and counts summarized using Rsubread [@liao2019]. This data is available from the Oldham Lab GitHub repository (https://github.com/oldhamlab/rnaseq.lf.hypoxia.molidustat). Differentially expressed transcripts were identified using DESeq2 [@love2014]. Gene set enrichment and transcription factor enrichment was performed using the fgsea and DoRothEA R packages, respectively [@korotkevich2021a; @garcia-alonso2019].

@@

(ref:m1) **Effects of hypoxia on extracellular metabolite fluxes in lung fibroblasts.** (**A**) Lung fibroblasts (LFs) were cultured in 21% or 0.5% oxygen beginning 24 h prior to time 0. Samples were collected every 24 h for 72 h. (**B**) Growth curves of LFs in each experimental condition (n = 8). (**C**) Growth rates from (B) were determined by robust linear modeling of log-transformed growth curves. (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to 21% oxygen at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to 21% oxygen treatment at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 8). By convention, negative fluxes indicate metabolite consumption. (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM (* p < 0.05).

(ref:m2) **Effects of pharmacologic prolyl hydroxylase inhibition on extracellular metabolite fluxes in lung fibroblasts.** (**A**) Lung fibroblasts (LFs) were treated with the prolyl hydroxlyase inhibitor molidustat (BAY, 10 μM) or DMSO beginning 24 h prior to time 0. Samples were collected every 24 h for 72 h. (**B**) Growth curves of LFs in each experimental condition (n = 8). (**C**) Growth rates from (B). (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to DMSO at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to DMSO at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 8). By convention, negative fluxes indicate metabolite consumption. (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM (* p < 0.05).

(ref:m3) **Metabolic flux analysis of lung fibroblasts following hypoxic and pharmacologic PHD inhibition.** (**A**) Mass isotopomer distribution (MID) of pyruvate (PYR) following 72 h labeling with [U-^13^C~6~]-glucose (GLC). (**B**) MID of citrate after 72 h labeling with [U-^13^C~6~]-GLC (**C**) MID of citrate after 72 h labeling with [U-^13^C~5~]-glutamine (GLN). Data are mean ± SEM (n = 4, p < 0.05 indicated as * 0.5% *v.* 21% oxygen, † BAY *v.* DMSO, ‡ Δoxygen *v.* ΔBAY). (**D**) Fraction of M5 citrate indicating reductive carboxylation after labeling with [U-^13^C~5~]-GLN in LFs and PASMCs (n = 3-4, * p < 0.05). (**E**) Ratio of modeled metabolic fluxes in 0.5% oxygen compared to 21% oxygen. Fluxes with non-overlapping confidence intervals are highlighted with arrows colored according to the magnitude of the fold change. Arrow thickness corresponds to the absolute flux measured in hypoxia. (**F**) Ratio of metabolic fluxes in BAY-treated cells compared to DMSO-treated control. Arrows are colored as in (E) and arrow weights correspond to the absolute flux as measured in BAY-treated cells.

(ref:m4) **PHD inhibition increases lactate uptake and oxidation.** (**A**) Exchange flux estimates for reductive carboxylation of α-ketoglutarate (AKG) to citrate (CIT). Data show the estimate with upper and lower bounds for LFs and PASMCs. (**B**) Exchange flux estimates for lactate import presented as in (A). (**C**) Mass isotopomer distributions of lactate (LAC), fructose-bisphosphate (FBP), pyruvate (PYR), and citrate (CIT) in LFs following 72 h labeling with [U-^13^C~3~] lactate. Data are mean ± SEM (n = 4, p < 0.05 indicated as * 0.5% *v.* 21% oxygen, † BAY *v.* DMSO, ‡ Δoxygen *v.* ΔBAY).

(ref:m5) **Metabolomic analysis of molidustat treatment in normoxia and hypoxia.** (**A**) LFs were cultured in standard growth medium and treated with molidustat (BAY, 10 μM) or vehicle (DMSO, 0.1%) in 21% or 0.5% oxygen conditions for 72 h. Growth rates were determined by linear fitting of log-transformed growth curves. (**B**) Glucose uptake in LFs treated with BAY cultured in hypoxia (note reversed *y*-axis). (**C**) Lactate efflux in LFs treated with BAY cultured in hypoxia. Data are mean ± SEM (* p < 0.05 BAY *v.* DMSO within a given oxygen exposure). (**D**) Principal component analysis of intracellular metabolites following 72 h of treatment as in (A) suggests a dominant effect of hypoxia over pharmacologic PHD inhibition on the metabolome (n = 7). (**E**) Volcano plot of intracellular metabolites of BAY-treated cells cultured in 21% or 0.5% oxygen. Circles indicate significantly increased (*blue*) and decreased (*red*) levels in hypoxia. (**F**) Metabolite set enrichment analysis of metabolites from (E). All KEGG pathways with p-values < 0.05 are shown. (**G**) Leading edge analysis of the TCA cycle metabolite set. Negative values indicate the relative enrichment associated with BAY treatment compared to hypoxia treatment. Abbreviations: PYR, pyruvate; SUC, succinate; PEP, phosphoenolpyruvate; CIT, citrate; AKG, α-ketoglutarate; MAL, malate; ACO, aconitate; FUM, fumarate. (**H**) NAD^+^ from LFs cultured as in (A). (**I**) NADH from LFs cultured as in (A). (**J**) NADH/NAD^+^ ratio from (H) and (I). NAD and NADH were determined by enzymatic assay. Data are mean ± SEM (n = 5, p < 0.05 with *black* * indicating a significant effect of treatment within a given oxygen tension and *colored* * indicating a significant effect of treatment within a given oxygen tension denoted by the color).

(ref:m6) **Transcriptomic analysis of molidustat treatment in normoxia and hypoxia.** (**A**) Principal component analysis of transcriptional changes in lung fibroblasts following 72 h of treatment with 0.5% oxygen or molidustat (BAY), separately or together (n = 4). (**B**) Volcano plot illustrating the effects of hypoxia in BAY-treated cells on gene expression. (**C**) Gene set enrichment analysis of transcripts from (B). (**D**) Volcano plot of those transcripts comprising the E2F Targets and G2/M Checkpoint Hallmark gene sets. (**E**) Volcano plot of those transcripts comprising the MYC Targets V1 and V2 Hallmark gene sets. (**F**) Volcano plot illustrating the results of a transcription factor enrichment analysis suggests mechanisms for differential regulation of gene expression following hypoxia or BAY treatment.

(ref:m7) **MYC regulates HIF-dependent glycolytic flux.** (**A**) Representative immunoblot of MYC protein expression in lung fibroblasts (LFs) following 72 h of treatment with 0.5% oxygen or molidustat (BAY, B). (**B**) Quantification of band densities from (A). (**C**) Working model illustrating that hypoxia-induced MYC antagonizes HIF-dependent metabolic and transcriptional events. (**D**) Representative immunoblot of LFs treated with siRNA targeting MYC (M) demonstrating adequate protein knockdown. (**E**) Growth rates of MYC-knockdown cells cultured in hypoxia. (**F**) Lactate efflux rates of MYC-knockdown cells cultured in hypoxia. (**G**) Representative immunoblot of LFs treated with MYC adenovirus. (**H**) Growth rates of MYC overexpressing cells cultured with BAY. (**I**) Lactate efflux rates of MYC overexpressing cells cultured with BAY. Data are mean ± SEM (n = 3-7, p < 0.05 as indicated by black * for differences within groups and colored * for differences between groups defined by the *x*-axis).

### 2022-08-02 manuscript/supplement.Rmd
finish edits prior to submission

(ref:s1) **Supporting data for extracellular flux calculations.** (**A**) Cell viability as assessed by live/dead cell staining with acridine orange plus propidium iodide staining did not differ between 21% and 0.5% oxygen culture conditions (n = 3 technical replicates). (**B**) Standard curve of lung fibroblast (LF) cell count *v.* total DNA by PicoGreen measurement used to interpolate cell numbers from DNA measurements. Data are mean ± SEM of three biological replicates. (**C**) Standard curve of PASMC cell count *v.* total DNA as in (B). (**D**) Total DNA measurements were compared to direct cell counts over the experimental time course. Cell counts and total DNA were obtained from the same sample wells. The slopes of the best-fit lines for 21% (*red*) and 0.5% (*blue*) samples were not statistically different. (**E**) Predicted well volumes were estimated from the change in culture plate mass over the experimental time course. Evaporation rates were different depending on the culture conditions and treatment. Although the mean evaporation rate is depicted, experiment-specific evaporation rates were used to calculate fluxes for each biological replicate (**F**) Metabolite accumulation (positive values) and degradation (negative values) rates. Data are mean ± SEM of 3-8 biological replicates. Rates significantly different from 0 (\*) based on a probability value < 0.05 using Student's one-sample *t*-test were incorporated into flux calculations.

(ref:s2) **Effects of hypoxia on extracellular metabolite fluxes in lung fibroblasts.** (**A**) Lung fibroblasts (LFs) were cultured in 21% (*red*) or 0.2% (*dark blue*) oxygen beginning 24 h prior to time 0. Samples were collected every 24 h for 72 h. (**B**) Growth curves of LFs in each experimental condition (n = 4). (**C**) Growth rates from (B) were determined by robust linear modeling of log-transformed growth curves. (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to 21% oxygen at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to 21% oxygen treatment at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 4). (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM. Comparisons were made using linear mixed effects models with oxygen as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (* p < 0.05).

(ref:s3) **Effects of hypoxia on extracellular metabolite fluxes in pulmonary artery smooth muscle cells.** (**A**) Pulmonary artery smooth muscle cells (PASMCs) were cultured in 21% (*red*) or 0.5% (*blue*) oxygen beginning 24 h prior to time 0. Samples were collected every 12 h for 48 h. (**B**) Growth curves of LFs in each experimental condition (n = 4). (**C**) Growth rates from (B) were determined by robust linear modeling of log-transformed growth curves. (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to 21% oxygen at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to 21% oxygen treatment at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 4). (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM. Comparisons were made using linear mixed effects models with oxygen as the fixed effect and biological replicate as a random effect with p-values adjusted using the multivariate *t* distribution (* p < 0.05).

@@

(ref:s7) **Isotopically non-stationary metabolic flux analysis.** (**A**) Metabolic flux model of LF metabolism in 21% oxygen. Arrows are colored by log~10~(flux). (**B**) Metabolic flux model of PASMC metabolism in 21% oxygen as in (A). (**C**) The ratio between PASMC and LF fluxes in normoxia were calculated. Fluxes with non-overlapping confidence intervals are highlighted with arrows colored according to the magnitude of the change. Arrow thickness corresponds to the absolute flux measured in LFs. (**D**) Ratio of metabolic fluxes in 0.5% oxygen compared to 21% oxygen in PASMCs. (**E**) LF fluxes were normalized to cell growth rate. Graph depicts the ratio of normalized metabolic fluxes in LFs cultured in 0.5% oxygen compared to 21% oxygen control. Fluxes with non-overlapping confidence intervals are highlighted to indicate significant changes.

>>>>

(ref:s1) **Supporting data for extracellular flux calculations.** (**A**) Cell viability as assessed by live/dead cell staining with acridine orange and propidium iodide did not differ between 21% and 0.5% oxygen culture conditions (n = 3 technical replicates). (**B**) Standard curve of lung fibroblast (LF) cell count *v.* total DNA by PicoGreen measurement used to interpolate cell numbers from DNA measurements. Data are mean ± SEM of three biological replicates. (**C**) Standard curve of PASMC cell count *v.* total DNA as in (B). (**D**) Total DNA measurements were compared to direct cell counts over the experimental time course. Cell counts and total DNA were obtained from the same sample wells. The slopes of the best-fit lines for 21% (*red*) and 0.5% (*blue*) samples were not statistically different. (**E**) Predicted well volumes were estimated from the change in culture plate mass over the experimental time course. Evaporation rates were different depending on the culture conditions and treatment. Although the mean evaporation rate is depicted, experiment-specific evaporation rates were used to calculate fluxes for each biological replicate (**F**) Metabolite accumulation (positive values) and degradation (negative values) rates. Data are mean ± SEM of 3-8 biological replicates. Rates significantly different from 0 (\*) based on a probability value < 0.05 using Student's one-sample *t*-test were incorporated into flux calculations.

(ref:s2) **Effects of hypoxia on extracellular metabolite fluxes in lung fibroblasts.** (**A**) Lung fibroblasts (LFs) were cultured in 21% or 0.2% oxygen beginning 24 h prior to time 0. Samples were collected every 24 h for 72 h. (**B**) Growth curves of LFs in each experimental condition (n = 4). (**C**) Growth rates from (B) were determined by robust linear modeling of log-transformed growth curves. (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to 21% oxygen at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to 21% oxygen treatment at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 4). By convention, negative fluxes indicate metabolite consumption. (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM (* p < 0.05).

(ref:s3) **Effects of hypoxia on extracellular metabolite fluxes in pulmonary artery smooth muscle cells.** (**A**) Pulmonary artery smooth muscle cells (PASMCs) were cultured in 21% or 0.5% oxygen beginning 24 h prior to time 0. Samples were collected every 12 h for 48 h. (**B**) Growth curves of LFs in each experimental condition (n = 4). (**C**) Growth rates from (B) were determined by robust linear modeling of log-transformed growth curves. (**D**) Representative immunoblot of LF protein lysates cultured as in (A). (**E**) Relative change in HIF-1α protein levels from (D) normalized to 21% oxygen at time 0 (n = 4). (**F**) Relative change in GLUT1 mRNA levels normalized to 21% oxygen treatment at time 0 (n = 4). (**G**) Relative change in LDHA mRNA levels as in (F). (**H**) Relative change in LDHA protein levels as in (E). (**I**) Extracellular fluxes of glucose (GLC) and lactate (LAC) (n = 4). By convention, negative fluxes indicate metabolite consumption. (**J**) Extracellular fluxes of pyruvate (PYR) and amino acids. Data are mean ± SEM (* p < 0.05).

@@

(ref:s7) **Isotopically non-stationary metabolic flux analysis.** (**A**) Metabolic flux model of LF metabolism in 21% oxygen. Arrows are colored by log~10~(flux). (**B**) Metabolic flux model of PASMC metabolism in 21% oxygen as in (A). (**C**) Ratio of metabolic fluxes in PASMCs compared to LFs. Fluxes with non-overlapping confidence intervals are highlighted with arrows colored according to the magnitude of the change. Arrow thickness corresponds to the absolute flux measured in LFs. (**D**) Ratio of metabolic fluxes in 0.5% oxygen compared to 21% oxygen in PASMCs. (**E**) LF fluxes were normalized to cell growth rate. Graph depicts the ratio of normalized metabolic fluxes in LFs cultured in 0.5% oxygen compared to 21% oxygen control. Fluxes with non-overlapping confidence intervals are highlighted to indicate significant changes.
